---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Kennedy"
author:  
  name: "Abby Kohan"
  email: "kohana@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---
16/20

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
freq_table <- table(Cars93$Type)
freq_table

rel_freq_table <- prop.table(freq_table)
rel_freq_table 
```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
freq_graph <- barplot(freq_table, main = "Frequency Bar Graph for Type", col = "lightgreen", xlab = "Type", ylab = "Count", cex.lab = 1.2, cex.axis = 1.2)
freq_graph 
```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
pie_chart <- pie(table(Cars93$Type), main = "Type", col=c(2,3,4,5,6,7), cex = 1.2)
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.
 What is the mode? (-1)

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
hp_hist <- hist(Cars93$Horsepower, col = "Red", xlab = "Horsepower", ylab = "Frequency", main = "Histogram of Horsepower", cex.lab = 1.2, cex.axis = 1.2)
hp_hist
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
hp_hist2 <- hist(Cars93$Horsepower, main = "New Histograom of Horsepower", xlab = "Horsepower", ylab = "Frequency", col = "skyblue", cex.lab = 1.2, cex.axis = 1.2, breaks = 25)
hp_hist2

## 25 Hp increments- each bin should have a width of 25. Not 25 bins. (-2)
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.
```{r}
#The histogram with breaks at 25 HP has more bins, providing a finer-grained view of the data. It can capture smaller variations and details within the distribution.
```


## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
```{r}
#Passengers is nominal and discrete. When making a histogram it could be difficult to visualize the data it masks certain categories such as cars with 1 and 3 that have no passengers meaning there should be no bar at 1 and 3.

pass_freq <- table(Cars93$Passengers)
pass_freq
hist(pass_freq, main = "Histogram of Passengers", xlab = "# of Passengers", ylab = "Frequency", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2, col = "yellow")
```

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)
```{r}
passengers <- c(Cars93$Passengers)
freq_table <- table(factor(passengers, levels = c("1", "2", "3", "4", "5", "6", "7", "8")))
#colors <- c("red", "blue", "orange")
barplot(freq_table, main = "Passenger Barplot", xlab = "# of Passengers", ylab = "Frequency", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2, col = "light yellow", ylim = c(0, max(freq_table) + 1))
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
length_freq <- table(Cars93$Length)
length_hist <- hist(length_freq, main = "Length Histogram", xlab = "Length", ylab = "Frequency", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2, col ="lightgreen")
length_hist
```

b) What is the shape of the distribution?
Skewed right

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)? Do they meet your expectations from part b? (-1)
```{r}
length_mean <- mean(Cars93$Length)
length_mean
length_median <- median(Cars93$Length)
length_median
```

You are done! ðŸ Don't forget to commit and push your .Rmd file to your Github repository before the due date.

